---
title: "Post DBEM data analysis for manuscript"
author: "William W.L. Cheung, Juliano Palacios Abrantes, Sarah Roberts"
date: "18/05/2022"
output: html_document
---

---
title: "pre_dbem"
author: "Sarah Roberts"
date: "5/4/2022"
output: html_document
---

```{r}
# For grid estimation and species selection
library(tidyverse)
library(sf)
library(viridis)
library(wesanderson)
library(worms)
library(readr)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

# For collaborative ease
library(here)

```



# Results

# Figure 1


```{r}
# MPA test grid
grids <- read_csv(here("./Data/mpa_scenarios/grid_m5point5.csv")) %>% 
  select(lon = coords.x1,
         lat = coords.x2,
         prop,
         status) %>% 
  mutate(cons_scen = 5) %>% 
  bind_rows(read_csv(here("./Data/mpa_scenarios/grid_30.csv")) %>% 
              select(lon = coords.x1,
                     lat = coords.x2,
                     prop,
                     status) %>% 
              mutate(cons_scen = 30)
  ) %>% 
  bind_rows(read_csv(here("./Data/mpa_scenarios/grid_15.csv")) %>% 
              select(lon = coords.x1,
                     lat = coords.x2,
                     prop,
                     status) %>% 
              mutate(cons_scen = 15)
  ) %>% 
  mutate(status = replace_na(status,"no protection"))



# test <- grids %>%
  # filter(status == "surrounding") %>%
  # mutate(bins = cut(prop, breaks = 3))

# Load worlkd map
World_map <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf"))
meow_sf <- st_read("../Data/spatial/MEOW/meow_ecos.shp") %>% 
  filter(REALM %in% c("Temperate Northern Atlantic","Arctic"))




  ggplot() +
    # geom_sf(data = World_map, aes()) +
    # geom_tile(data = subset(grids, !is.na(bins)),
    geom_tile(data = grids,
    aes(
      x = lon,
      y = lat,
      fill = status,
      color = status
    ),
    alpha = 0.7
  ) +
    geom_sf(dat = meow_sf, aes(), color = "black", fill =NA) +
    geom_sf(data = World_map, aes(), fill = "grey90") +
    scale_fill_manual("Conservation status",values = c(
      "white",
      wes_palette("Zissou1")[5],
      "darkblue"
      # wes_palette("Rushmore1")[3]
    )
                      ) +
    scale_color_manual("Conservation status",values = c(
      "white",
      wes_palette("Zissou1")[5],
      "darkblue"
      # wes_palette("Rushmore1")[3]
      
                                 )
                      ) +
    # scale_fill_viridis("Conservation status", discrete = T, alpha = 0.7,option = "D",direction = -1) +
    # scale_color_viridis("Conservation status",discrete = T, alpha = 0.7,option = "D",direction = -1) +
    theme_bw() +
    labs(x = "Longitude", y = "Latitude") +
    theme(legend.position = "") +
    facet_wrap(~cons_scen) +
    coord_sf(xlim = c(-42.5,69),
             ylim = c(35,90)
             ) +
    ggsave("../Figures/scenarios_a_ch.png",
           width = 10,
           height = 5
           )
```



Double check the DBEM is doing what we want it to do

```{r read_data}

# MPA test grid
grid_30 <- read_csv(here("./Data/mpa_scenarios/grid_30.csv"))

mpa_grid <- grid_30 %>% 
  select(index = gridID,status,prop)

unique(mpa_grid$status)

# DBEM lat lon
coords <- read.csv(here("./Data/Lon_Lat_DBEM.txt"),header = F) %>% 
  select(index = V1,
         lon = V2,
         lat = V3)


# MCP data
mcp_data <- read.table("/Volumes/Enterprise/Data/northsea_mpa_sust_fish/Raw/C6IPSLtest/600069/600069Catch2020.txt", quote="\"", comment.char="")
colnames(mcp_data) <- c("index","mcp")

# ABD data
abd_data <- read.table("/Volumes/Enterprise/Data/northsea_mpa_sust_fish/Raw/C6IPSLtest/600069/600069Abd2020.txt", quote="\"", comment.char="")
colnames(abd_data) <- c("index","abd")

# Join both

test_data <-
  mcp_data %>% 
  full_join(abd_data) %>% 
  gather("var","value",mcp:abd)





```

```{r}

test_data %>% 
  left_join(coords) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = value
    )
  ) +
  facet_wrap(~var)

```



```{r}

# Compare catch between surrounding and open ocean
# Surrounding should be more, at least in average
test_data %>% 
  left_join(mpa_grid) %>% 
  mutate(value = ifelse(is.na(value),0,value),
         status = ifelse(is.na(status),"open",status)
         ) %>% 
  group_by(status,var) %>% 
  summarise(
    sum = sum(value, na.rm = T),
    mean = mean(value, na.rm = T),
    n = n()
  )




```



```{r check_protected_areas}

test_data %>% 
  left_join(mpa_grid) %>% 
  mutate(value = ifelse(is.na(value),0,value),
         status = ifelse(is.na(status),"open",status)
         ) %>% 
  filter(status == "protected")




```

```{r map_me}

map_data <- mpa_grid %>% 
  left_join(coords,
            by = "index") %>% 
  left_join(test_data,
            by = "index") %>%
  mutate(status = ifelse(is.na(status),"open",status)) %>% 
  filter(!is.na(value))


World_map <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf"))

ggplot() +
  geom_sf(data = World_map, aes()) +
  geom_sf(data = fao_regions, aes()) +
  geom_tile(data = map_data,
    aes(
      x = lon,
      y = lat,
      fill = value,
      color = value
    )
  )


```



# Table S1

*Can you please get the number of fishes and invertebrates, and create a table that list the species as a supplementary table? Here is the final list of species:* https://drive.google.com/open?id=1xlsMJGGVTIax9GQXoMVOx8EgFC1eNgbU&authuser=wwlcheung%40gmail.com&usp=drive_fs


```{r}


nea_spplist <- read_csv("../Data/species/nea_spplist.csv")
final_spplist <- read_csv("../Data/species/final_spplist.csv")

# Get info necessary
ssplit <- nea_spplist %>% 
  filter(taxon_key %in% final_spplist$x)

# Cerate Table S1 using worms
table_s1 <- worms::wormsbynames(ssplit$taxon_name) %>% 
  mutate(vert_invert = ifelse(phylum == "Chordata","vertebrate","invertebrate")) %>% 
  select(taxon_name=valid_name,kingdom:genus,vert_invert) %>% 
  left_join(ssplit,
            by = "taxon_name") %>% 
  select(taxon_name,common_name,kingdom:genus.x,vert_invert)

# Write table for paper
# write_csv(table_s1,"../Data/table_s1.csv")

# Estimate numbers for intext
table_s1 %>% 
  group_by(vert_invert) %>% 
  tally()
   



```

# Sensitivity analysis

We created 3 different MPA configurations for each conservation scenario and then ran the DBEM for 10 species (Table S1) with different life history characteristics (e.g. pelagics, demersals, bivalves, gastropod, cephalopods, etc.). THe sensitivity analysis was done using the GFDL following SSP 585

1.- Pick species with wide ranging life history characteristics: small pelagic, medium pelagic, large pelagic, small demersal, medium demersal, large demersal, inverts (bivalues/gastropod), cephalopods, sharks/rays (long-lived), a bathydemersal (deep sea). 

## Species selection

```{r species_selection}

# manually select speices for sensitivity analysis
sens_taxa <- c(600115, # Atlantic bonito
               601357, # European sprat
               604329, # spotted ray
               690273, # Norway lobster
               600069, # Atlantic cod
               601381, # Haddock,
               600898, # Blue shark
               690088, # common octopus
               690166, # Grooved carpet shell
               605094  # Blackbellied angler
               )


# Get spp list table
sens_spplist <- read_csv("../Data/species/final_spplist.csv") %>% 
  dplyr::rename(taxon_key = x) %>% 
  left_join(read_csv("../Data/species/nea_spplist.csv"),
            by = "taxon_key") %>% 
  filter(taxon_key %in% sens_taxa)

# Save for reference
write_csv(sens_spplist, "../Data/species/sens_spplist.csv")

# Save for DBEM
sens_spplist %>% 
  select(taxon_key) %>% 
  write.table(., file="../Data/species/NsSppList19.txt",
              sep="\t", 
              col.names = F, 
              row.names = F)

sens_spplist

```












# Bugs

## Test F075 bug

**Fixed**

```{r}

library("tidyverse")
library("here")

Lon_Lat_DBEM <- read.csv("Data/Lon_Lat_DBEM.txt", header=F)
colnames(Lon_Lat_DBEM) <- c("index","lon","lat")

grid_mpa <- read_csv(here("./Data/mpa_scenarios/grid_30.txt")) %>% 
  select(index = gridID, zone = zone.x,status,prop)


grid_mpa <- read_csv(here("./Data/mpa_scenarios/grid_15.csv")) %>% 
  select(index = gridID, zone = zone.x,status,prop)



load("~/Desktop/600054Catch.RData")


abd <- as.data.frame(sppabdfnl) 
colnames(abd) <- seq(1951,2100,1)

abd <- abd %>% 
  bind_cols(coords) %>% 
  gather("year",'value',1:150) %>% 
  filter(!is.na(value))


# Linear value
grid_mpa %>% 
  left_join(abd,
            by = "index") %>% 
  # filter(status != "open") %>%
  # View()
  # mutate(status = ifelse(is.na(status),"open",status)) %>% 
  # View()
  group_by(year,status) %>% 
  summarise(m = mean(value,na.rm=T)) %>%
  # View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = log10(m),
      color = status
    )
  )

```

