---
title: "Post DBEM data analysis for manuscript"
author: "William W.L. Cheung, Juliano Palacios Abrantes, Sarah Roberts"
date: "18/05/2022"
output: html_document
---


```{r}
# For grid estimation and species selection
library(tidyverse)
library(sf)
library(viridis)
library(wesanderson)
library(worms)
library(readr)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

# For collaborative ease
library(here)

```



# Results

## Figure 1


```{r}
# MPA test grid
grids <- read_csv(here("./Data/mpa_scenarios/grid_m5point5.csv")) %>% 
  select(lon = coords.x1,
         lat = coords.x2,
         prop,
         status) %>% 
  mutate(cons_scen = 5) %>% 
  bind_rows(read_csv(here("./Data/mpa_scenarios/grid_30.csv")) %>% 
              select(lon = coords.x1,
                     lat = coords.x2,
                     prop,
                     status) %>% 
              mutate(cons_scen = 30)
  ) %>% 
  bind_rows(read_csv(here("./Data/mpa_scenarios/grid_15.csv")) %>% 
              select(lon = coords.x1,
                     lat = coords.x2,
                     prop,
                     status) %>% 
              mutate(cons_scen = 15)
  ) %>% 
  mutate(status = replace_na(status,"no protection"))



# test <- grids %>%
  # filter(status == "surrounding") %>%
  # mutate(bins = cut(prop, breaks = 3))

# Load worlkd map
World_map <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf"))
meow_sf <- st_read("../Data/spatial/MEOW/meow_ecos.shp") %>% 
  filter(REALM %in% c("Temperate Northern Atlantic","Arctic"))




  ggplot() +
    # geom_sf(data = World_map, aes()) +
    # geom_tile(data = subset(grids, !is.na(bins)),
    geom_tile(data = grids,
    aes(
      x = lon,
      y = lat,
      fill = status,
      color = status
    ),
    alpha = 0.7
  ) +
    geom_sf(dat = meow_sf, aes(), color = "black", fill =NA) +
    geom_sf(data = World_map, aes(), fill = "grey90") +
    scale_fill_manual("Conservation status",values = c(
      "white",
      wes_palette("Zissou1")[5],
      "darkblue"
      # wes_palette("Rushmore1")[3]
    )
                      ) +
    scale_color_manual("Conservation status",values = c(
      "white",
      wes_palette("Zissou1")[5],
      "darkblue"
      # wes_palette("Rushmore1")[3]
      
                                 )
                      ) +
    # scale_fill_viridis("Conservation status", discrete = T, alpha = 0.7,option = "D",direction = -1) +
    # scale_color_viridis("Conservation status",discrete = T, alpha = 0.7,option = "D",direction = -1) +
    theme_bw() +
    labs(x = "Longitude", y = "Latitude") +
    theme(legend.position = "") +
    facet_wrap(~cons_scen) +
    coord_sf(xlim = c(-42.5,69),
             ylim = c(35,90)
             ) +
    ggsave("../Figures/scenarios_a_ch.png",
           width = 10,
           height = 5
           )
```



Double check the DBEM is doing what we want it to do

```{r read_data}

# MPA test grid
grid_30 <- read_csv(here("./Data/mpa_scenarios/grid_30.csv"))

mpa_grid <- grid_30 %>% 
  select(index = gridID,status,prop)

unique(mpa_grid$status)

# DBEM lat lon
coords <- read.csv(here("./Data/Lon_Lat_DBEM.txt"),header = F) %>% 
  select(index = V1,
         lon = V2,
         lat = V3)


# MCP data
mcp_data <- read.table("/Volumes/Enterprise/Data/northsea_mpa_sust_fish/Raw/C6IPSLtest/600069/600069Catch2020.txt", quote="\"", comment.char="")
colnames(mcp_data) <- c("index","mcp")

# ABD data
abd_data <- read.table("/Volumes/Enterprise/Data/northsea_mpa_sust_fish/Raw/C6IPSLtest/600069/600069Abd2020.txt", quote="\"", comment.char="")
colnames(abd_data) <- c("index","abd")

# Join both

test_data <-
  mcp_data %>% 
  full_join(abd_data) %>% 
  gather("var","value",mcp:abd)





```

```{r}

test_data %>% 
  left_join(coords) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = value
    )
  ) +
  facet_wrap(~var)

```



```{r}

# Compare catch between surrounding and open ocean
# Surrounding should be more, at least in average
test_data %>% 
  left_join(mpa_grid) %>% 
  mutate(value = ifelse(is.na(value),0,value),
         status = ifelse(is.na(status),"open",status)
         ) %>% 
  group_by(status,var) %>% 
  summarise(
    sum = sum(value, na.rm = T),
    mean = mean(value, na.rm = T),
    n = n()
  )




```



```{r check_protected_areas}

test_data %>% 
  left_join(mpa_grid) %>% 
  mutate(value = ifelse(is.na(value),0,value),
         status = ifelse(is.na(status),"open",status)
         ) %>% 
  filter(status == "protected")




```

```{r map_me}

map_data <- mpa_grid %>% 
  left_join(coords,
            by = "index") %>% 
  left_join(test_data,
            by = "index") %>%
  mutate(status = ifelse(is.na(status),"open",status)) %>% 
  filter(!is.na(value))


World_map <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf"))

ggplot() +
  geom_sf(data = World_map, aes()) +
  geom_sf(data = fao_regions, aes()) +
  geom_tile(data = map_data,
    aes(
      x = lon,
      y = lat,
      fill = value,
      color = value
    )
  )


```



## Table S1

*Can you please get the number of fishes and invertebrates, and create a table that list the species as a supplementary table? Here is the final list of species:* https://drive.google.com/open?id=1xlsMJGGVTIax9GQXoMVOx8EgFC1eNgbU&authuser=wwlcheung%40gmail.com&usp=drive_fs


```{r}


nea_spplist <- read_csv("../Data/species/nea_spplist.csv")
final_spplist <- read_csv("../Data/species/final_spplist.csv")

# Get info necessary
ssplit <- nea_spplist %>% 
  filter(taxon_key %in% final_spplist$x)

# Cerate Table S1 using worms
table_s1 <- worms::wormsbynames(ssplit$taxon_name) %>% 
  mutate(vert_invert = ifelse(phylum == "Chordata","vertebrate","invertebrate")) %>% 
  select(taxon_name=valid_name,kingdom:genus,vert_invert) %>% 
  left_join(ssplit,
            by = "taxon_name") %>% 
  select(taxon_name,common_name,kingdom:genus.x,vert_invert)

# Write table for paper
# write_csv(table_s1,"../Data/table_s1.csv")

# Estimate numbers for intext
table_s1 %>% 
  group_by(vert_invert) %>% 
  tally()
   



```

# Sensitivity analysis

We ran a sensitivity analysis to explore the influence of the random MPA allocation to the biomass projection for each conservation scenario. We created 3 different MPA configurations for each conservation scenario and then ran the DBEM for 10 species (Table S1) with different life history characteristics (e.g. pelagics, demersals, bivalves, gastropod, cephalopods, etc.) using the GFDL following SSP 585.

To test for the effect of different configurations we first aggregated the yearly biomass per species withing each grid cell status (protected, surrounding and unprotected) and then computed the yearly mean +- 2*sd of the three extra configurations. Finally, we explored how far from the variation the configuration used in the main analysis.

Results from the sensitivity analysis show that most of the runs used for the present analysis are withing the variation of the different configurations. In few cases the projected biomass of the main configuration is beyond the variation. Nevertheless, the projected biomass is quite closed to the variation border which at the same time is quite narrow (e.g., Atlantic bonito and cod for protected cells of the 5% scenario).




## Species selection

```{r species_selection}

# manually select speices for sensitivity analysis
sens_taxa <- c(600115, # Atlantic bonito
               601357, # European sprat
               604329, # spotted ray
               690273, # Norway lobster
               600069, # Atlantic cod
               601381, # Haddock,
               600898, # Blue shark
               690088, # common octopus
               690166, # Grooved carpet shell
               605094  # Blackbellied angler
               )


# Get spp list table
sens_spplist <- read_csv("../Data/species/final_spplist.csv") %>% 
  dplyr::rename(taxon_key = x) %>% 
  left_join(read_csv("../Data/species/nea_spplist.csv"),
            by = "taxon_key") %>% 
  filter(taxon_key %in% sens_taxa)

# Save for reference
write_csv(sens_spplist, "../Data/species/sens_spplist.csv")

# Save for DBEM
sens_spplist %>% 
  select(taxon_key) %>% 
  write.table(., file="../Data/species/NsSppList19.txt",
              sep="\t", 
              col.names = F, 
              row.names = F)

sens_spplist

```


## MPA scenarios

```{r}

sen_mpa_grids <- list.files("../Data/mpa_scenarios",full.names = T,pattern = ".csv")

x = sen_mpa_grids[1]

# Make function

my_read <- function(x){
  
  if(str_detect(x,"point") == T){
    sen <- paste0(0,str_sub(x,35,37))
  }else{
  sen <- str_sub(x,28,31)
  }
  out <- read_csv(x) %>% 
    mutate(sen = sen)
  
  return(out)
  
}

# Load data
all_mpa_grids <- bind_rows(
  lapply(sen_mpa_grids, my_read)
) %>% 
  select(sen,
         lon = coords.x1,
         lat = coords.x2,
         prop,status) %>% 
  mutate(prop = replace_na(0),
         status = ifelse(is.na(status),"unprotected",status)
         )

unique(all_mpa_grids$sen)


# Make map
ggplot() +
    # geom_sf(data = World_map, aes()) +
    # geom_tile(data = subset(grids, !is.na(bins)),
    geom_tile(data = all_mpa_grids,
    aes(
      x = lon,
      y = lat,
      fill = status,
      color = status
    ),
    alpha = 0.7
  ) +
  geom_sf(dat = meow_sf, aes(), color = "black", fill =NA) +
    geom_sf(data = World_map, aes(), fill = "grey90") +
    scale_fill_manual("Conservation status",values = c(
      wes_palette("Zissou1")[5],
      "darkblue",
      "white"
    )
                      ) +
    scale_color_manual("Conservation status",values = c(
      wes_palette("Zissou1")[5],
      "darkblue",
      "white"
      
                                 )
                      ) +
    theme_bw() +
    labs(x = "Longitude", y = "Latitude") +
    theme(legend.position = "right",
          legend.key = element_rect(colour = 'black', fill = 'white', linetype='solid')
          ) +
    facet_wrap(~sen) +
    coord_sf(xlim = c(-42.5,69),
             ylim = c(35,90)
             ) +
  ggsave("../Figures/sens_scenarios.png",
           width = 10,
           height = 5
           )



```


## Sensitivity Analisys 


```{r sens_protocol, eval = F}

# Get scenario list
scenarios <- list.files("/Users/juliano/Data/northsea_mpa_sust_fish/sens",full.names = T)

# get selected species
species <- read.csv("../Data/species/sens_spplist.csv") %>% 
  pull(taxon_key)

# Years of projected data
years = seq(1951,2100,1)

# For testing
# rm(final_scen,final_yr,partial_yr,final_spp)
# j = 1
# i = 1
# y = 1


sens_fx <- function(specie){
  print(specie)
  
  # Loop trough all scenarios
  for(i in 1:length(scenarios)){

    Load scenario grids
      grid_df <- read.csv(paste0("../Data/mpa_scenarios/grid_",str_sub(scenarios[i],62,65),".csv"))
      grid <- grid_df  %>% 
        select(V1 = gridID,
               status
               ) %>% 
        mutate(status = ifelse(is.na(status),"unprotected",status))
      
    # Loop trough all years
    for(y in 1:length(years)){
      
      # Read data
      to_read <- paste0(scenarios[i],"/",specie,"/",specie,"Abd",years[y],".txt")
      partial_yr <- read.table(to_read, quote="\"", comment.char="") %>% 
          mutate(year = years[y])
      
      
      # Combine year data
      if(y == 1){
        final_yr <- partial_yr
      }else{
        final_yr <- bind_rows(final_yr,partial_yr)
      }
      
    } # Close years
      
      # Aggregate abd by year
      agg_yr <- final_yr %>% 
        left_join(grid,
                  by = "V1") %>% 
        filter(!is.na(status)) %>% 
        group_by(status,year) %>%
        summarise(total_abd = sum(V2,na.rm = T), .groups = "keep") %>% 
        mutate(specie = specie,
               scen = str_sub(scenarios[i],59,63),
               run = paste0("run",str_sub(scenarios[i],64,65))
               )
               
      # Compile all scenario data
      if(i == 1){
        final_scen <- agg_yr
      }else{
        final_scen <- bind_rows(final_scen,agg_yr)
      }
  } # Close scenarios    
      
  return(final_scen)
} # Close function

# sens_fx(specie = species[5]) %>% View()

# scens_df <- bind_rows(
  # lapply(species,sens_fx)
# )

# Save dataset for future computation
 # readr::write_csv(scens_df, "../Data/scensitivity_df.csv")

```

## Double check

```{r}

scens_df <- read_csv("../Data/scensitivity_df.csv") %>% 
  rename(taxon_key = specie) %>% 
  left_join(read_csv("../Data/species/nea_spplist.csv"),
            by = "taxon_key")


unique(scens_df$scen)
unique(scens_df$year)
unique(scens_df$specie)
unique(scens_df$run)
unique(scens_df$status)


```


## Plot it

### Option for totals

```{r}

ggplot(scens_df) +
  geom_line(
    aes(
      x = year,
      # y = total_abd/1000000,
      y = log10(total_abd),
      color = run,
      linetype = status
    )
  ) +
  facet_wrap(scen~specie, scales = "free_y",ncol = 10) +
  theme(legend.position = "top")

```


```{r}

# sens_spplist

averages <- scens_df %>% 
  filter(run != "run") %>% 
  group_by(taxon_key,common_name,status,scen,year) %>% 
  summarise(mean = mean(total_abd),
            sd = sd(total_abd)
            )


ggplot(averages) +
  geom_line(
    aes(
      x = year,
      y = mean,
      color = status
    )
  ) +
  geom_ribbon(data = averages,
              aes(
                x = year,
                ymin = mean-(2*sd),
                ymax = mean+(2*sd),
                fill = status
              ),
              alpha = 0.5
              ) +
  geom_point(data = scens_df %>% filter(run == "run"),
    aes(
      x = year,
      y = total_abd
    ),
    color = "grey",
    size = 0.1
  ) +
  ylab("Projected biomass") +
  xlab("Year") +
  facet_wrap(scen~common_name, scales = "free_y",ncol = 5) +
  theme(legend.position = "top") +
  theme_classic()

ggsave("sensitivity_analsyis.png",
       last_plot(),
       width = 15,
       height = 10)


```



### Statistics


```{r}

scens_df %>% 
  group_by(common_name,)





```


# Bugs

## Test F075 bug

**Fixed**

```{r}

library("tidyverse")
library("here")

Lon_Lat_DBEM <- read.csv("Data/Lon_Lat_DBEM.txt", header=F)
colnames(Lon_Lat_DBEM) <- c("index","lon","lat")

grid_mpa <- read_csv(here("./Data/mpa_scenarios/grid_30.txt")) %>% 
  select(index = gridID, zone = zone.x,status,prop)


grid_mpa <- read_csv(here("./Data/mpa_scenarios/grid_15.csv")) %>% 
  select(index = gridID, zone = zone.x,status,prop)



load("~/Desktop/600054Catch.RData")


abd <- as.data.frame(sppabdfnl) 
colnames(abd) <- seq(1951,2100,1)

abd <- abd %>% 
  bind_cols(coords) %>% 
  gather("year",'value',1:150) %>% 
  filter(!is.na(value))


# Linear value
grid_mpa %>% 
  left_join(abd,
            by = "index") %>% 
  # filter(status != "open") %>%
  # View()
  # mutate(status = ifelse(is.na(status),"open",status)) %>% 
  # View()
  group_by(year,status) %>% 
  summarise(m = mean(value,na.rm=T)) %>%
  # View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = log10(m),
      color = status
    )
  )

```

