---
title: "pre_dbem"
author: "Sarah Roberts"
date: "5/4/2022"
output: html_document
---

```{r}
# For grid estimation and species selection
library(spatialEco)
library(ggplot2)
library(tidyverse)
library(geosphere)
library(raster)
library(units)
library(matrixStats)

library(sf)
library(tidyverse)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

# For collaborative ease
library(here)


```


# MPA grids

for dbem grid that is overlapping FAO27 randomly select 30% points from those 30 points figure out which points are surrounding them (within .75 dec.degrees to get the diagonal)

do the same for the 15% (from the selected 30 - take 50%)

```{r}
fao_27 <- read_sf(here("./Data/spatial/Zone_27.shp"))


grid <- read.csv(here("./Data/Lon_Lat_DBEM.txt"), header = FALSE)
colnames(grid) <- c("gridID", "long", "lat")
grid_sp <- st_as_sf(grid, coords = c("long", "lat"), crs = st_crs(fao_27))


sf::sf_use_s2(FALSE)
grid_intersect_27 <- point.in.poly(grid_sp, fao_27)

grid_intersect_27_df <- as.data.frame(grid_intersect_27) 

grid_intersect_27_df <- grid_intersect_27_df[!is.na(grid_intersect_27_df$zone), ] #just zone 27
rm(grid, grid_intersect_27, grid_sp)

ggplot() + 
  geom_sf(data = fao_27) + 
  geom_point(data = grid_intersect_27_df, aes(x = coords.x1, y = coords.x2), size = .2) 
`%!in%` <- Negate(`%in%`)

grid_30 <- grid_intersect_27_df %>% sample_frac(.3)
no_30 <- grid_intersect_27_df  %>% dplyr::filter(gridID %!in% c(grid_30$gridID) )

grid_15 <- grid_30 %>% sample_frac(.5)
no_15 <- grid_intersect_27_df  %>% dplyr::filter(gridID %!in% c(grid_15$gridID) )

grid_30_sp <- st_as_sf(grid_30, coords = c("coords.x1", "coords.x2"), crs = NA)
no_30_sp <- st_as_sf(no_30, coords = c("coords.x1", "coords.x2"), crs = NA)

grid_15_sp <- st_as_sf(grid_15, coords = c("coords.x1", "coords.x2"), crs = NA)
no_15_sp <- st_as_sf(no_15, coords = c("coords.x1", "coords.x2"), crs = NA)
grid_orig <- st_as_sf(grid_intersect_27_df, coords = c("coords.x1", "coords.x2"), crs = NA) #by leaving crs as NA it will calculate distance in decimal degrees



#now we need to get the touching grid cells surrounding the 30 and 15 grid. These will be within .75 dec.degrees 

```

## Distance - 30 
```{r}
dist_30 <- st_distance(no_30_sp, grid_30_sp)
colnames(dist_30) <- grid_30$gridID #column names are the protected gridIDs

mins <- rowMins(dist_30, na.rm = T)
dist_30_n <- as.data.frame(mins)
dist_30_n$assoc_ID <- apply(dist_30, 1, function(x) colnames(dist_30)[which.min(x)]) #find which one they are closest to - need to know for proportion calc later on

dist_30_n$gridID <- no_30$gridID

n <- 1111398*.75 #1 degree is that many meters (approx) - we want anything 
#this is really big to subset, lets get out smallest distance first 

dist_30_touching <- subset(dist_30_n, dist_30_n$mins < .75)
dist_30_touching <-left_join(dist_30_touching, no_30, by = "gridID")
dist_30_touching <- dist_30_touching %>% add_count(assoc_ID) %>% mutate("prop" = (1/n) + 1)

ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_30, aes(x = coords.x1, y = coords.x2), size = .2, colour = "red") + 
  geom_point(data = dist_30_touching, aes(x = coords.x1, y = coords.x2), size = .2, colour = "blue") + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

#ookay that seemed to work 
```

#### Clean up the dataset 

```{r}
dist_30_touching$status <- "surrounding"
dist_30_touching <- dist_30_touching %>% dplyr::select(-mins, -n, -assoc_ID)
grid_30$prop <- 0
grid_30$status <- "protected"

pt_30 <- rbind(dist_30_touching, grid_30)

grid_30_final <- left_join(grid_intersect_27_df, pt_30, by = c("gridID","coords.x1", "coords.x2"))


#checkit 
ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_30_final, aes(x = coords.x1, y = coords.x2, colour = status), size = .2)  + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

write.csv(grid_30_final, here("./Data/mpa_scenarios/grid_30.csv"))
```


### Distance - 15 
```{r}
dist_15 <- st_distance(no_15_sp, grid_15_sp)
colnames(dist_15) <- grid_15$gridID #column names are the protected gridIDs

mins <- rowMins(dist_15, na.rm = T)
dist_15_n <- as.data.frame(mins)
dist_15_n$assoc_ID <- apply(dist_15, 1, function(x) colnames(dist_15)[which.min(x)]) #find which one they are closest to - need to know for proportion calc later on

dist_15_n$gridID <- no_15$gridID

n <- 1111398*.75 #1 degree is that many meters (approx) - we want anything 
#this is really big to subset, lets get out smallest distance first 

dist_15_touching <- subset(dist_15_n, dist_15_n$mins < .75)
dist_15_touching <-left_join(dist_15_touching, no_15, by = "gridID")
dist_15_touching <- dist_15_touching %>% add_count(assoc_ID) %>% mutate("prop" = (1/n) + 1)

ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_15, aes(x = coords.x1, y = coords.x2), size = .2, colour = "red") + 
  geom_point(data = dist_15_touching, aes(x = coords.x1, y = coords.x2), size = .2, colour = "blue") + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

#ookay that seemed to work 
```

#### Clean up the dataset 

```{r}
dist_15_touching$status <- "surrounding"
dist_15_touching <- dist_15_touching %>% dplyr::select(-mins, -n, -assoc_ID)
grid_15$prop <- 0
grid_15$status <- "protected"

pt_15 <- rbind(dist_15_touching, grid_15)

grid_15_final <- left_join(grid_intersect_27_df, pt_15, by = c("gridID","coords.x1", "coords.x2"))


#checkit 
ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_15_final, aes(x = coords.x1, y = coords.x2, colour = status), size = .2)  + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

write.csv(grid_15_final, here("./Data/mpa_scenarios/grid_15.csv"))
```


## Add in MPA data 

Note - I found these areas in ArcMap - the MPA shapefile was too big to work with in R - so I selected those that intersect FAO 27, dissolved to one shapefile that includes Marine and Coastal, projected to albers equal area, then calculated the area (in meters). Projected the FAO data into albers equal area as well. I can put those in github, but because they are shapefiles they are kind of big.  

coastal mpa area (meters)- 24315764487.799999 + 93245644267.800003
marine mpa area - 819401638008

FAO area - 14635120094800

so 5.5% of the area is stricly marine protected 819401638008/14635120094800
and 6.4 percent of the area is coastal and marine protected 24315764487.799999 + 93245644267.800003 + 819401638008/14635120094800

I'll do them both to have them both - but we can decide what to use. 

6.4 percent (wich is .2133% of 30 (6.4/30))

```{r}
grid_cm <- grid_30 %>% sample_frac(.2133)
no_cm <- grid_intersect_27_df  %>% dplyr::filter(gridID %!in% c(grid_cm$gridID) )

grid_cm_sp <- st_as_sf(grid_cm, coords = c("coords.x1", "coords.x2"), crs = NA)
no_cm_sp <- st_as_sf(no_cm, coords = c("coords.x1", "coords.x2"), crs = NA)
grid_orig <- st_as_sf(grid_intersect_27_df, coords = c("coords.x1", "coords.x2"), crs = NA) #by leaving crs as NA it will calculate distance in decimal degrees




```

### Distance - 6.4 percent (wich is .2133% of 30 (6.4/30))
```{r}
dist_cm <- st_distance(no_cm_sp, grid_cm_sp)

colnames(dist_cm) <- grid_cm$gridID #column names are the protected gridIDs

mins <- rowMins(dist_cm, na.rm = T)
dist_cm_n <- as.data.frame(mins)
dist_cm_n$assoc_ID <- apply(dist_cm, 1, function(x) colnames(dist_cm)[which.min(x)]) #find which one they are closest to - need to know for proportion calc later on

dist_cm_n$gridID <- no_cm$gridID

n <- 1111398*.75 #1 degree is that many meters (approx) - we want anything 
#this is really big to subset, lets get out smallest distance first 

dist_cm_touching <- subset(dist_cm_n, dist_cm_n$mins < .75)
dist_cm_touching <-left_join(dist_cm_touching, no_cm, by = "gridID")
dist_cm_touching <- dist_cm_touching %>% add_count(assoc_ID) %>% mutate("prop" = (1/n) + 1)

ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_cm, aes(x = coords.x1, y = coords.x2), size = .2, colour = "red") + 
  geom_point(data = dist_cm_touching, aes(x = coords.x1, y = coords.x2), size = .2, colour = "blue") + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

#ookay that seemed to work 
```

#### Clean up the dataset 

```{r}
dist_cm_touching$status <- "surrounding"
dist_cm_touching <- dist_cm_touching %>% dplyr::select(-mins, -n, -assoc_ID)
grid_cm$prop <- 0
grid_cm$status <- "protected"

pt_cm <- rbind(dist_cm_touching, grid_cm)

grid_cm_final <- left_join(grid_intersect_27_df, pt_cm, by = c("gridID","coords.x1", "coords.x2"))


#checkit 
ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_cm_final, aes(x = coords.x1, y = coords.x2, colour = status), size = .2)  + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

write.csv(grid_cm_final, here("../Data/mpa_scenarios/grid_cm6point4.csv"))
```

### 5.5 percent (wich is .1833% of 30 (5.5/30))
```{r}
grid_m <- grid_30 %>% sample_frac(.1833)
no_m <- grid_intersect_27_df  %>% dplyr::filter(gridID %!in% c(grid_m$gridID) )

grid_m_sp <- st_as_sf(grid_m, coords = c("coords.x1", "coords.x2"), crs = NA)
no_m_sp <- st_as_sf(no_m, coords = c("coords.x1", "coords.x2"), crs = NA)
grid_orig <- st_as_sf(grid_intersect_27_df, coords = c("coords.x1", "coords.x2"), crs = NA) #by leaving crs as NA it will calculate distance in decimal degrees



```

### Distance - 6.4 percent (wich is .2133% of 30 (6.4/30))
```{r}
dist_m <- st_distance(no_m_sp, grid_m_sp)
colnames(dist_m) <- grid_m$gridID #column names are the protected gridIDs

mins <- rowMins(dist_m, na.rm = T)
dist_m_n <- as.data.frame(mins)
dist_m_n$assoc_ID <- apply(dist_m, 1, function(x) colnames(dist_m)[which.min(x)]) #find which one they are closest to - need to know for proportion calc later on

dist_m_n$gridID <- no_m$gridID

n <- 1111398*.75 #1 degree is that many meters (approx) - we want anything 
#this is really big to subset, lets get out smallest distance first 

dist_m_touching <- subset(dist_m_n, dist_m_n$mins < .75)
dist_m_touching <-left_join(dist_m_touching, no_m, by = "gridID")
dist_m_touching <- dist_m_touching %>% add_count(assoc_ID) %>% mutate("prop" = (1/n) + 1)

ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_m, aes(x = coords.x1, y = coords.x2), size = .2, colour = "red") + 
  geom_point(data = dist_m_touching, aes(x = coords.x1, y = coords.x2), size = .2, colour = "blue") + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

#ookay that seemed to work 
```

#### Clean up the dataset 

```{r}
dist_m_touching$status <- "surrounding"
dist_m_touching <- dist_m_touching %>% dplyr::select(-mins, -n, -assoc_ID)
grid_m$prop <- 0
grid_m$status <- "protected"

pt_m <- rbind(dist_m_touching, grid_m)

grid_m_final <- left_join(grid_intersect_27_df, pt_m, by = c("gridID","coords.x1", "coords.x2"))


#checkit 
ggplot() + 
  geom_sf(data = fao_27) +
  geom_point(data = grid_m_final, aes(x = coords.x1, y = coords.x2, colour = status), size = .2)  + 
  coord_sf(xlim=c(-35, -30), ylim=c(40,45), expand = TRUE)

write.csv(grid_m_final, here("../Data/mpa_scenarios/grid_m5point5.csv"))
```


# Species selection

Here we will select the species to model based on the FAO region 27

## Get FAO area

```{r}

# read FAO region 27
# FAO regions
fao_regions <- my_sf("FAO") %>% 
  filter(F_LEVEL == "MAJOR",
         F_CODE == 27)
st_crs(fao_regions) = 4326
  
ggplot(fao_regions) + 
  geom_sf()

```

## Get Spp distributions

```{r}

# Read DBEM coordinates (full)
dbem_coords <- my_path("Spa", "DBEM", name = "complete_lon_lat_dbem.csv", read = T, header = T)

# Create function to crop distribution to north sea
crop_dist <- function(taxonkey){
  
  
  # Reads distribution and crops it to FAO
output <- read_distribution(taxonkey, output = "data", root_path = "/Volumes/DATA") %>% 
  bind_cols(dbem_coords) %>% 
  filter(dis >0) %>% 
  filter(f_code == 27)

# Detect stock
if(nrow(output)>0){
  return(data.frame(TaxonKey = taxonkey))
  }else{
    return(data.frame(TaxonKey = NA))
  }
  
}


# Read spp list

# Read species list
SppTaxonName <- read.csv("/Volumes/DATA/JULIANO_NEYMAR/EmergingFish/Data/SAU/SppTaxonName.csv")

# Select species
spp_list <- bind_rows(lapply(SppTaxonName$TaxonKey, crop_dist)) %>% 
  left_join(SppTaxonName) %>% 
  filter(!is.na(TaxonKey)) %>% 
  janitor::clean_names()

write_csv(spp_list,"../Data/nea_spplist.csv")

```


# Re-allocation of *f_mort*

When a grid is closed to fishing (i.e., a MPA), the fishing mortality of that grid cell is distributed to the inmmediate surrounding gridcells. This way, the model accounts for the re-distribution of fishing effort that ofetn happens when an MPA is stabished.


## Get Data

```{r reallocation_data, eval = T, echo = F}

# Make select function dplyr
select <- dplyr::select

# MPA test grid
grid_30 <- read_csv(here("./Data/mpa_scenarios/grid_30.csv"))

mpa_grid <- grid_30 %>% 
  select(index = gridID,status)

# DBEM lat lon
coords <- read.csv(here("./Data/Lon_Lat_DBEM.txt"),header = F) %>% 
  select(index = V1,
         lon = V2,
         lat = V3)


```

## Design method

```{r}
# test
coords %>% 
  mutate(fmort = runif(259200,0,1)) %>%  # Simulate DBEM's fmort
  # Re estimate effort
  left_join(mpa_grid,
            by = "index") %>% 
  mutate(
    status = ifelse(is.na(status),"open",status),
    prop = ifelse(status == "protected",0,
                  ifelse(status == "surrounding",runif(3,1,2),
                         1)
                  ),
    fmort_prop = fmort * prop
    ) %>% 
  filter(status == "surrounding") %>% 
  View()

# Ok

```

Sounds like the way we can do it is to proportionally divide the protected cell into the surrounding ones. This way, $if status = protected, prop == 0$, consequently, if there are 4 cells surrounding the protected one, then $prop =1.25$, if there are 2 surrounding cells then $prop =1.5$, if there are 10 surrounding cells then $prop =1.1$, and so on... Finally, if $status = no protection$ then $prop = 1$.

With that information we can then re-estimate fishing mortality ($f_{mort}$) (i.e., re-allocate fishing effort) in the surrounding cells as

$$\hat{f}_{mort} = f_{mort} * prop$$
where $\hat{f}_{mort}$ is the fishing mortality considering the re-allocation fo fishing effort due to the closure if fishing within the protected grid cell. This means that, in a surrounding cell where, lets say $f_{mort}$ = 0.2 and prop = 1.25, the actual fishing mortality (i.e., $\hat{f}_{mort}$) is actually 0.25 ($\hat{f}_{mort} = 0.25$). Subsequently, in cases where $prop = 0$ (i.e., protected grid cells) and $f_{mort}$ > 0, $\hat{f}_{mort}$ = 0. Finally, if the cell is not protected nor surrounding a protected grid cell then  $\hat{f}_{mort} = f_{mort}$, that is, fishing mortality will be the one estimated by the DBEM assuming no effects from the MPA

With this in mind, we need a (new) `grid_mpa` dataset that looks something like the one below so we can input that to the DBEM

```{r grid_map_new}
# Final table
# I think this is how it might work out

 mpa_grid %>% 
  full_join(coords) %>% 
  mutate(
    status = ifelse(is.na(status),"open",status),
    # Randomly simulate proportions for exemplification
    prop = ifelse(status == "protected",0,
                  ifelse(status == "surrounding",runif(3,1,2),
                         1)
                  )
    ) %>%
  arrange(index) %>% 
  select(prop)

```

